Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-07 paper.tex
more refinements, mostly sections 4,5,6,7

\subsection{Our acception of anthropomorphism and related terminology}

@@

\paragraph{Pre-interaction phase\\}
\hl{needs to be written, \eg people form expectations \etc and this leads to the ICA (\ICA).}

\paragraph{Initial Capital of Anthropomorphism\\}
\label{sec:ica}

@@

\paragraph{The Novelty Effect\\}
\label{sec:noveltyeffect}

Novelty effects have been described with interactive technologies in general \citep{rogers_diffusion_1995} and also in several long-term HRI studies in workplaces \citep{huttenrauch_fetch-and-carry_2003,mutlu_robots_2008}, in schools and public places \citep{gockley_designing_2005,kanda_interactive_2004,kanda_communication_2005}, or elder care centers \citep{sabelli_conversational_2011}. Further, also with domestic robots it has been found that people's interest, engagement, and fascination with the robot decreased over time \citep{fink_living_2013,fernaeus_how_2010,sung_robots_2009}.

To illustrate how the novelty of a robot can evoke anthropomorphic effects, we
present an example from our long-term ethnographic study with Roomba vacuum
cleaning robots in households \citep{fink_living_2013}. The observed scenario describes how an elderly
woman (71 years, living on her own with a little dog in a small row house) was
setting up and using the robot for the first time during about 10 minutes.

The woman (E.) is reading aloud parts of the ``quick start manual'': \emph{``When
Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
intensely in that area.''} She laughs and turns to the robot, talking directly to
it \emph{``Today you can light up darling!''}. E. carries the Roomba to the entrance
door and puts it on the floor [floor type: tiles, partially covered with little
Persian carpets], she explains \emph{``It has to start from here, this is where
you enter the house!''}. After pressing the CLEAN button, the robot quickly gets
stuck with the carpet in the entrance hall, consequently it reports error 1 [
two-tone \textit{``uh-oh''} sound followed by a narrated voice that starts off with \textit{``error
1''}]. E. is very surprised \emph{``What, it is even talking to me?''}, she lifts the
robot off and puts it back for cleaning. At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{``Nice!''} or \emph{``Good job!''} or \emph{``Bravo!''}. Once, when the Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{``There you forgot something, what about over
there? Look, there you still have something to clean up!''}, and she points at
where she wants the Roomba to go. Further, she uses deictic gestures and tells the
robot: \emph{``No, go this way! Come over here!''}. The woman keeps observing the
robot and wonders about its intentions, asking it directly: \emph{``Why do you
leave the kitchen? There is still something there on the fringe!''}. E. turns to
the experimenter, and tells her: \emph{``She [the robot] loves the hall!''}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender (\emph{``darling''}
/ \emph{``she''}); praising the robot (\emph{``Good job!''}); using pointing
gestures toward the robot, or trying to direct it around with commands; attributing
a personal preference / intention to the robot (\emph{``she loves the hall!''}).

During the following visits at the woman's place, E. did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. The observed
effects were likely due to the novelty of the robot, and the fact that E. was not familiar with the robot and its behavior.
As \cite{persson_anthropomorphism_2000} names it: \textit{``Anthropomorphism is a way of making sense of complex behavior in the world around us.''}. Thus, as soon as this complexity is reduced, also anthropomorphism might be reduced.

@@

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
potentially facing an agent aiming at ``human-level intelligence'' (to take
McCarthy's words). This is a transient state that quickly vanishes, and the
projected anthropomorphism then starts to decrease while the human observes and
recognizes that the behavior of the robot is generally \emph{predictable} and
possibly \emph{deceptive} (for instance, it talks but does not understand when
we talk ; it has eyes but does not recognize everyday objects we show ; etc.).

@@

By \emph{disruptive behaviors}, we mean any behavior exhibited by the robot that
is unexpected by the user: for instance, a robot may usually follow always the
same route to go to two different places in a house, and suddenly change the
route. The actual underlying reason may span from a bug to the detection of a
new possible obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}.

@@

\fixme{JF: Severin, we should at some point mention what the ``expectation mismatch'' is ... I haven't found that so far ... but I am also not sure about how to explain this. I think that, the we already took into account the user's expectations in the human-factor.}

@@

We also discuss two specific aspects: the cognitive impact of disruptive
behaviors (section~\ref{sec:disruptive}) on mutual modelling, and how our models
relate to the \emph{in-the-moment} versus \emph{reflective} perspectives on agency in HRI~\citet{takayama_perspectives_2012}.

@@

We propose three different cognitive stages (Figure~\ref{fig:cognitivemodel}),
which are related but do not directly match the three phases of
the model of the dynamics of anthropomorphism. Note that these three stages are not mutually exclusive, they may overlap in time.

@@

The \emph{stage 1} is the instinctive, pre-cognitive identification of living
peers. Studies conducted by~\citet{rosenthal-vonderputten_experimental_2013},
who investigated the neural correlates of emotional reactions of humans towards
a robot, supports the idea that humans tend to anthropomorphize robots
intuitively in this pre-cognitive way. {\it Empathy} is characteristic of this
stage~\citep{rosenthalvonderPutten2013neural}.  It is also at this stage that
automatic activation of the human's mirror neurons system  in response to
viewing the robots action and human automatic emotional responses are
contributing at a lower level to the intuitive anthropomorphization (see Section
2.2.2 \fixme{put ref}).

The intuitive anthropomorphization in the cognitive stage 1 might be characteristic of what \cite{takayama_perspectives_2012} calls the \textit{in-the-moment} perspective on agency with a robot compared to the \textit{reflective} perspective (see paragraph ``artifact-centered perspective'' in Section \ref{sec:anthropomorphism}). 
For instance, we reflectively
might \textit{not} perceive agency in a social robot, but it can feel quite differently
in the moment of interaction. \cite{takayama_perspectives_2012}
argues that neglecting to separate reflective perspectives from in-the-moment
perspectives of agency is one of the major sources of confusion when people talk
and write about anthropomorphism. There seems to be a disconnection between what
people consciously perceive and how they respond to stimuli that they may not
consciously perceive. That means, in the initial phase of interacting with a
novel device \textit{(initialization)} (see Figure \ref{fig:dynamics}) people might not respond consciously but
rather mindlessly \citep{nass_machines_2000}, and in turn anthropomorphize more.
Only after some time of
\textit{familiarization}, they might respond in a more reflective manner (in turn, the tendency to anthropomorphize might decrease). This
can be illustrated by the fact that people tend to deny interacting with
computational systems as if they were people and yet they respond to computers
in many ways that are remarkably similar to how they respond to people
\citep{reeves_media_1996}. \cite{takayama_perspectives_2012} also applies a
cognitive viewpoint on the two different perspectives of in-the-moment \vs
reflective to illustrate differences. She states that in-the-moment perceptions
of agency are largely shaped by bottom-up perceptual processes, evoking very
immediate responses (\eg to the cues emitted by the artifact's design (``artifact-centered perspective''). In contrast, according to Takayama,
reflective perceptions are more often shaped by top-down processes because of
the nature of reflective thought.

@@

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human reaches the cognitive \emph{stage 2}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human first
projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We
hypothesize that the nature of the projected mental model, as well as how deep
the human engages in this projection, might be driven by the same parameters as
we mentioned for the \emph{initial capital of anthropomorphism}
(section~\ref{sec:ica}).  At this stage the human might also change his attitude
towards the robot, paying more attention to social cue than to low-level
behaviour which may reinforce the resonnance mechanisms.

The cognitive \emph{stage 3} is reached after a \emph{contextualized} interaction.
A \emph{contextualized} interaction is \emph{explicitly purposeful} (the purpose
of the interaction, be it purely entertainment, is explicit and conscious to the
human), and takes place in an environment that fosters a stronger cognitive (and
possibly affective/social) commitment from the human in the interaction
(typically, at home). During this interaction, the human iteratively restates
and reshapes his/her behavioral and mental model of the robot (\emph{How does
the robot react to such and such situation/input?  What does the robot know
about me? About itself? About our environment? What can the robot learn?}, etc.).

@@

\subsection{Relation to the dynamics of anthropomorphism}

These cognitive stages overlap over time but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism. In particular,
the cognitive stages 1 and 2 are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive stage 2
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

The cognitive stage 3 also overlaps with the \emph{Familiarization} phase: as
the human gets used to the robot, he/she restates and adapts its
cognitive model of the robot by iteratively reshaping pre-existent, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation process
depicted in the cognitive stage 3 reached a stable state, \ie the user's
experience with the robot is correctly supported by the cognitive model he/she
has built.

@@

Case IIIa and IIIb correspond to misinterpretations of the robot behavior. Case
IIIb may actually lead to an increased level of anthropomorphism since the human
will (wrongfully) attribute intentionality to the robot, while case IIIa is
expected to lead to a lower anthropomorphism level.  However, in those two
cases, the next occurrence of an expected behavior, if correctly interpreted
(case I or II), is likely to lead to stronger effects due to a larger delta
between expectations and actual observed behavior.

@@

\fixme{Discuss here the general findings of the paper: the global shape of the
model of the dynamics, ... // JF: yep, I agree, we should also come back to our initial hypotheses}

@@

With \textit{time} as a factor, we
referred to long-term interaction with the robot, which is related to what has been
described as \textit{novelty effect} but also accounts for the user getting used
to the robot.

As outlined before, the tendency to anthropomorphize is motivated by
a person's wish to make sense of an agent which might be difficult to understand
(in its functionality or behavior, for instance). In turn, a user might ascribe
intentions or emotions to the system because the system's output was unexpected
for the user. Based on this explanation, we estimate that when the user has
familiarized herself / himself with the system, thus reached the point of when
the system is usable and explainable, the tendency to anthropomorphize
decreases.

However, there exist views that suggest just the opposite, namely, that familiarization with an agent \textit{increases} the tendency to anthropomorphize \citep{eddy_attribution_1993}. Similarly, \cite{duffy_anthropomorphism_2003} reports that several psychological experiments and human-robot interaction experiments showed that familiarity may also ease social acceptance and even tend to \textit{increase} people's tendency to anthropomorphize.

@@

One needs to keep in mind that anthropomorphism arises in an interaction. We believe that as such, anthropomorphism can be better understood when studied and measured in a real, direct human-robot interaction compared to a mediated interaction. In a direct interaction scenario one could also take the human's behavior into account, as it seems to reflect an anthropomorphic interaction \citep{krach_can_2008,hegel_understanding_2008}.
Of course, the scenario and measurements used depend on the research questions, so a direct human-robot interaction might not always be preferred.

@@

Regarding the second issue, how and with which means anthropomorphism is measured, the central point is that ``human-likeness'' itself is complex and manifold. It as a quite diverse and abstract concept, and instead of asking such a vague question, whether a robot is human-like or not, \cite{kahn_jr._robotic_2006} suggest to ask for more concrete constructs that are typical or unique of the concept of ``human-likeness''.
For instance, \cite{ruijten_introducing_2014} propose a 25-item questions to measure various concrete aspects of human-likeness. Consequently, anthropomorphism can be modeled along a scale that accounts for lower and higher levels of it.

Still, we also feel that measuring anthropomorphism (or social engagement with a robot) is challenging because the phenomenon is partially
unconscious, its mediation is largely subject-dependent\fxnote{Should we put
here different interaction profiles from the Ranger study?}, and we found the
phenomenon generally difficult to verbalize for users.

@@

The main tools for
qualitative measurement of anthropomorphic effect are questionnaires and semi-structured interviews. We have
explored both rating-scales in a questionnaire, and open questions in an interview (inspired from \citep{kahn_jr._robotic_2006,weiss_i_2009}).
In the Domino study, interviews were transcribed and participant's answers to the key questions that were assessing an anthropomorphic projection onto the robot, were analyzed. For instance, one question to see how far children would ascribe moral standing to the robot was \emph{``If you had this robot at home, and you were to leave for 2 weeks of holidays, would it be ok to leave the robot alone at home?''}. A point for anthropomorphism was counted, when the participant's answer indicated an anthropomorphic view of the robot. This process led to a ``qualitative score'' that reflects how far the robot is anthropomorphized. We used the following constructs to measure how far the robot is perceived as a human-like agent:

@@

\subsection{Sustaining long-term interaction}

One main challenge that robots with applications in daily life contexts face is to sustain long-term interaction.

\paragraph{Impact of the behavioral complexity of robots}

To remain tractable, most experiments, and even more when they investigate
long-term interactions, are carried out with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism \AntMax is
reached at the end of the novelty effect, and then decreases. A natural question
is whether we can design robots such as \AntMax remains high, possibly higher
than $\ant[t_{novelty~peak}]$ (note that to know if sustaining strong
anthropomorphic interaction is desirable is entirely another question).

Intuitively, the behavioral complexity of the robot, and in particular, the ability
for the robot to evolve over time and adapt its behavior to the user as the user
familiarizes his/herself with the robot (\ie, \emph{co-familiarize}), seems to play a
key role.

\fixme{JF: Severin, I am not sure what you want to say here.}

\subsection{Ethical aspects of anthropomorphism and human-like form of robots}

@@

After all discussion, there remains the question how far a sustained anthropomorphic (human social) interaction between a human and a non-human can be favored, regarding ethical aspects. We need to rethink carefully the consequences that emotional bonds between humans and robots can have. They might be beneficial on one hand but can also hold negative implications. How far should a robot's design encourage us to lastingly project feelings or intentions to it? Why should a machine even react to us as if it had emotions? Can this be the reality we want to live in? What happens as soon as the artificial agent is considered a \textit{moral agent} \cite{sullins_when_2006}? Also, the use of \textit{socially interactive robots} in 
contexts involving elderly care or in rehabilitation and with autistic children is a controversially discussed topic 
\cite{robins_robots_2005}. Do we want robots to comfort us; robots that make us eventually even falling in love with them?
The anthropomorphic design of robots is critical. Why do we even need this? The HRI / robotics community needs to discuss these and other questions.

@@

Our data suggest that there are qualitative differences and dynamics over time
in how humans relate to different types of robots. More concretely, we found
that people's perception of human-likeness in the robot and in turn their
human-social engagement with the robot was not only based on how much the robot
resembled a human (mostly in its behavioral cues) but that this was determined
by multiple factors, including individual characteristics of the human user but
also the context of use, for instance. Interestingly, our findings also suggest
that how far people anthropomorphize a robot is likely to change over time with
growing experience of the user with the system.

Based on these observations and an extensive literature review, we would like to
propose a model that accounts for variances and dynamics in anthropomorphism.
This model of the \textit{``Dynamics of Anthropomorphism''} considers several
impact factors, including the human user, the anthropomorphic design of the
non-human agent, the context of use, and the interaction experience (time). With
this initial model, we would like to extend the current understanding of
anthropomorphism in robotics and human-robot interaction. This might in turn
lead to a discussion about the short- and long-term benefits and other effects
of the human-like (anthropomorphic) design of robots and other artificial
agents.

@@

While anthropomorphism is traditionally empirically understood as the
interactions between the anthropomorphic design of a robot and the psychological
determinants of the user, it appears that the duration and context of the
interaction are also key factors, and that anthropomorphism needs to be
understood as a dynamic phenomenon. In this article, we propose a new formal
model of anthropomorphism that accounts for these three factors and also
explicits the dynamics of anthropomorphism. We introduce the concepts of
\emph{initial} and \emph{stabilized levels of anthropomorphism} as compound
factors to characterize the profile of a given anthropomorphic interaction.

The article also discusses the cognitive correlates of anthropomorphism. We
proposes to identify three cognitive stages corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We show how these stages relate to observable anthropomorphic
effects, and how they evolve over time.

@@

>>>>

\subsection{Terminology used in this article}

@@

\paragraph{Pre-interaction phase and Initial Capital of Anthropomorphism (ICA)\\}
\label{sec:ica}

@@

%\paragraph{The Novelty Effect\\}
%\label{sec:noveltyeffect}

We assume that the short phase of initialization is strongly affected by the so-called novelty effect.
Novelty effects have been described with interactive technologies in general \citep{rogers_diffusion_1995} and also in several long-term HRI studies in workplaces \citep{huttenrauch_fetch-and-carry_2003,mutlu_robots_2008}, in schools and public places \citep{gockley_designing_2005,kanda_interactive_2004,kanda_communication_2005}, or elder care centers \citep{sabelli_conversational_2011}. Further, also with domestic robots it has been found that people's interest, engagement, and fascination with the robot decreased over time \citep{fink_living_2013,fernaeus_how_2010,sung_robots_2009}.

To illustrate how the novelty of a robot can evoke anthropomorphic effects, we
present an example from our long-term ethnographic study with Roomba vacuum
cleaning robots in households \citep{fink_living_2013}. The observed scenario describes how an elderly
woman (E., 71 years, living on her own with a little dog in a small row house) was
setting up and using the robot for the first time during about 10 minutes.

%The woman (E.) is reading aloud parts of the ``quick start manual'': \emph{``When
%Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
%intensely in that area.''} She laughs and turns to the robot, talking directly to
%it \emph{``Today you can light up darling!''}. E. carries the Roomba to the entrance
%door and puts it on the floor [floor type: tiles, partially covered with little
%Persian carpets], she explains \emph{``It has to start from here, this is where
%you enter the house!''}. After pressing the CLEAN button, the robot quickly gets
%stuck with the carpet in the entrance hall, consequently it reports error 1 [
%two-tone \textit{``uh-oh''} sound followed by a narrated voice that starts off with \textit{``error
%1''}]. E. is very surprised \emph{``What, it is even talking to me?''}, she lifts the
%robot off and puts it back for cleaning.
At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{``Nice!''} or \emph{``Good job!''} or \emph{``Bravo!''}. Once, when the Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{``There you forgot something, what about over
there? Look, there you still have something to clean up!''}, and she points at
where she wants the Roomba to go. Further, she uses deictic gestures and tells the
robot: \emph{``No, go this way! Come over here!''}. The woman keeps observing the
robot and wonders about its intentions, asking it directly: \emph{``Why do you
leave the kitchen? There is still something there on the fringe!''}. E. turns to
the experimenter, and tells her: \emph{``She [the robot] loves the hall!''}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender, praising the robot, using pointing
gestures toward the robot, or trying to direct it around with commands, and attributing
a personal preference / intention to the robot (\emph{``she loves the hall!''}).

During the following visits at the woman's place, E. did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. The observed
effects were likely due to the novelty of the robot, and the fact that E. was not familiar with the robot and its behavior.
As \cite{persson_anthropomorphism_2000} names it: \textit{``Anthropomorphism is a way of making sense of complex behavior in the world around us.''} Thus, as soon as this complexity is reduced, also anthropomorphism might be reduced.

@@

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
potentially facing an agent aiming at ``human-level intelligence'' (to take
McCarthy's words). This is a transient state that quickly vanishes, and the
projected anthropomorphism then starts to decrease while the human observes and
recognizes that the behavior of the robot is generally \emph{predictable} and
possibly \emph{deceptive} (for instance, it talks but does not understand when
we talk; it has eyes but does not recognize everyday objects we show; \etc).

@@

By \emph{disruptive behaviors}, we mean any behavior exhibited by the robot that
appears unexpected to the user: for instance, a robot may usually follow always the
same route to go from one place in a house to the other, but suddenly it might change the
route. The actual underlying reason may span from a bug to the detection of a
new obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}. (We will come back to the role of disruptive behaviors later, in Section \ref{sec:disruptive}.)

@@

%\fixme{JF: Severin, we should at some point mention what the ``expectation mismatch'' is ... I haven't found that so far ... but I am also not sure about how to explain this. I think that, the we already took into account the user's expectations in the human-factor.}

@@

We also discuss two specific aspects: the cognitive impact of disruptive
behaviors (section~\ref{sec:disruptive}) on mutual modeling, and how our models
relate to the \emph{in-the-moment} versus \emph{reflective} perspectives on agency in HRI~\citep{takayama_perspectives_2012}.

@@

We propose three different cognitive stages (Figure~\ref{fig:cognitivemodel}), which are not mutually exclusive but overlap over time. Note that these three cognitive stages are related but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism.

@@

The \emph{Stage 1} is the instinctive, pre-cognitive identification of living
peers. Studies conducted by~\citet{rosenthal-vonderputten_experimental_2013},
who investigated the neural correlates of emotional reactions of humans towards
a robot, supports the idea that humans tend to anthropomorphize robots
intuitively in this pre-cognitive way. {\it Empathy} is characteristic of this
stage~\citep{rosenthalvonderPutten2013neural}.  It is also at this stage that
automatic activation of the human's mirror neurons system  in response to
viewing the robots action and human automatic emotional responses are
contributing at a lower level to the intuitive anthropomorphization.

The intuitive anthropomorphization in the cognitive Stage 1 might be characteristic of what \cite{takayama_perspectives_2012} calls the \textit{in-the-moment} perspective on agency with a robot compared to the \textit{reflective} perspective (see paragraph ``artifact-centered perspective'' in Section \ref{sec:anthropomorphism}). 
For instance, we reflectively
might \textit{not} perceive agency in a social robot, but it can feel quite differently
in the moment of interaction. \cite{takayama_perspectives_2012}
argues that neglecting to separate reflective perspectives from in-the-moment
perspectives of agency is one of the major sources of confusion when people talk
and write about anthropomorphism. There seems to be a disconnection between what
people consciously perceive and how they respond to stimuli that they may not
consciously perceive. That means, in the initial phase of interacting with a
novel device \textit{(initialization)} (see Figure \ref{fig:dynamics}) people might not respond consciously but
rather mindlessly \citep{nass_machines_2000}, and in turn anthropomorphize more.
Only after some time of
\textit{familiarization}, they might respond in a more reflective manner (in turn, the tendency to anthropomorphize might decrease). This
can be illustrated by the fact that people tend to deny interacting with
computational systems as if they were people and yet they respond to computers
in many ways that are remarkably similar to how they respond to people
\citep{reeves_media_1996}. \cite{takayama_perspectives_2012} also applies a
cognitive viewpoint on the two different perspectives of in-the-moment \vs
reflective to illustrate differences. She states that in-the-moment perceptions
of agency are largely shaped by bottom-up perceptual processes, evoking very
immediate responses (\eg to the cues emitted by the artifact's design (``artifact-centered perspective''). In contrast, according to Takayama,
reflective perceptions are more often shaped by top-down processes because of
the nature of reflective thought.

@@

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human reaches the cognitive \emph{Stage 2}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human first
projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We
hypothesize that the nature of the projected mental model, as well as how deep
the human engages in this projection, might be driven by the same parameters as
we mentioned for the \emph{initial capital of anthropomorphism}
(section~\ref{sec:ica}).  At this stage the human might also change his attitude
towards the robot, paying more attention to social cue than to low-level
behavior which may reinforce the resonance mechanisms.

The cognitive \emph{Stage 3} is reached after a \emph{contextualized} interaction.
A \emph{contextualized} interaction is \emph{explicitly purposeful} (the purpose
of the interaction, be it purely entertainment, is explicit and conscious to the
human), and takes place in an environment that fosters a stronger cognitive (and
possibly affective/social) commitment from the human in the interaction
(typically, at home). During this interaction, the human iteratively restates
and reshapes his/her behavioral and mental model of the robot (\emph{How does
the robot react to such and such situation/input?  What does the robot know
about me? About itself? About our environment? What can the robot learn?}, etc.).

@@

\paragraph{Relation to the dynamics of anthropomorphism\\}

As mentioned before, the three cognitive stages proposed here do not match with the three phases in the \emph{Dynamics of Anthropomorphism}.
In particular,
the cognitive Stages 1 and 2 are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive Stage 2
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

The cognitive Stage 3 also overlaps with the \emph{Familiarization} phase: as
the human gets used to the robot, he/she restates and adapts its
cognitive model of the robot by iteratively reshaping pre-existent, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation process
depicted in the cognitive Stage 3 reached a stable state, \ie the user's
experience with the robot is correctly supported by the cognitive model he/she
has built.

Another way to think of the three proposed cognitive stages is to relate them to the \textbf{user experience} which is characteristic of these three stages. For instance, \cite{norman_emotional_2003} characterizes  three user experience dimensions (related to emotional attachment), which occur when people use or see a product for the first time: \emph{(1)} the \emph{visceral level}, which is the first impression of a product based on its experience; at this level people do not think about a product but make spontaneous judgments; \emph{(2)} the \emph{behavioral level}, in which people use and experience the product, appraise its functions, and consider aspects such as usefulness and usability; \emph{(3)} the \emph{reflective level}, in which consciousness takes part in the process, and past experiences are taken into account.
In a study which was carried out in a shopping mall, \cite{weiss_i_2009} explored the emotional attachment during first time reactions of children and adults with the robotic dog AIBO. The authors also argue that it is important to understand the long-term development of people's social bonds to robots.

@@

Case IIIa and IIIb correspond to misinterpretations of the robot behavior. Case
IIIb may actually lead to an increased level of anthropomorphism since the human
will (wrongfully) attribute intentionality to the robot, while case IIIa is
expected to lead to a lower anthropomorphism level.  However, in those two
cases, the next occurrence of an expected behavior, if correctly interpreted
(case I or II), is likely to lead to stronger effects due to a larger $\delta$
between expectations and actual observed behavior.

@@

Coming back to our two main hypotheses, on one hand we investigated how far anthropomorphism is based on multiple factors, namely, the design of the robot, the personal characteristics of the human user, and the context / situation in which the robot is used.
We elaborated this hypothesis by reviewing literature from various related disciplines. Based on the outcomes of this, on the other hand, we proposed a model of the \emph{Dynamics of anthropomorphism}. We offered a cognitive interpretation of this model, as well as illustrations from our own previous work.
However, our model has limitations, for instance, regarding the general effect of time and familiarization on anthropomorphism. We will discuss this in the following section.

@@

With \textit{time} as a factor, we
referred to long-term interaction with the robot, which is related to what has been
described as \textit{novelty effect} but also accounts for the user getting used
to the robot. Several studies described that after some time users interacted less with the robot or lost interest in it. Sustaining long-term interaction is one of the main challenges that robots with applications in daily life contexts face. Our question is that, if we would like overcome the novelty effect, how far does the robot's behavioral complexity play a role? Should the robot vary its behavior and behave unexpectedly from time to time?

Most experiments, and even more when they investigate
long-term interactions, are carried out with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism \AntMax is
reached with the novelty effect, and then decreases. A natural question
is whether we can design robots such that \AntMax remains high, possibly higher
than $\ant[t_{novelty~peak}]$ (note that to know if sustaining strong
anthropomorphic interaction is desirable is entirely another question, see Section \ref{sec:ethics}).

Intuitively, the behavioral complexity of the robot, and in particular, the ability
for the robot to evolve over time and \emph{adapt its behavior to the user} as the user
familiarizes his/herself with the robot (\ie, \emph{co-familiarize}), seems to play a
key role.

Our model is based on the assumption that anthropomorphism tends to decrease with the user becoming more familiar to the robot.
We explain this by the general motivation to anthropomorphize an agent which might be difficult to understand
(in its functionality or behavior, for instance). In turn, a user might ascribe
intentions or emotions to the system because the system's output was unexpected
for the user, however he/she still wants to make sense of it. Consequently, we estimate that when the user has
familiarized herself/himself with the system, thus reached the point of when
the system is usable and explainable, the tendency to anthropomorphize
decreases.

However, there exist views that suggest just the opposite, namely, that familiarization and more interaction with an agent \textit{increase} the tendency to anthropomorphize \citep{eddy_attribution_1993}. Similarly, \cite{duffy_anthropomorphism_2003} reports that several psychological experiments and human-robot interaction experiments showed that familiarity may also ease social acceptance and even tend to \textit{increase} people's tendency to anthropomorphize.

@@

One needs to keep in mind that anthropomorphism arises in an interaction. We believe that as such, anthropomorphism can be better understood when studied and measured in a real, direct human-robot interaction compared to a mediated interaction. In a direct interaction scenario one could also take the human's behavior into account, as it seems to reflect an anthropomorphic interaction \citep{krach_can_2008,hegel_understanding_2008,weiss_i_2009}.
Of course, the scenario and measurements one uses depend on the research questions, so a direct human-robot interaction might not always be preferred.

@@

Regarding the second issue, how and with which means anthropomorphism is measured, one of the view existing tools to do so is 5-item anthropomorphism part of the Godspeed questionnaire \citep{bartneck_measurement_2008}. On 5 point semantic differential scales, people are asked to rate the following constructs: fake - natural, machinelike - humanlike, unconscious - conscious, artificial - lifelike, moving rigidly - moving elegantly. The critical point that we see with this is that ``human-likeness'' itself is complex and manifold. Being such an abstract concept, instead of asking the vague question, whether a robot is human-like or not, \cite{kahn_jr._robotic_2006} suggest to ask for more concrete constructs that are typical or unique of the concept of ``human-likeness''.
For instance, \cite{ruijten_introducing_2014} propose a 25-item questions to measure various concrete aspects of human-likeness. Consequently, anthropomorphism can be modeled along a scale that accounts for lower and higher levels of it.

Still, we also feel that measuring anthropomorphism (or social engagement with a robot) is challenging because the phenomenon is partially
unconscious, its mediation is largely subject-dependent, and we found the
phenomenon generally difficult to verbalize for users.

@@

The main tools for
qualitative measurement of anthropomorphic effect are questionnaires and semi-structured interviews. We have
explored both rating-scales in a questionnaire, and open questions in an interview.
In the Domino study, interviews were transcribed and participant's answers to the key questions that were assessing an anthropomorphic projection onto the robot, were analyzed. For instance, one question to see how far children would ascribe moral standing to the robot was \emph{``If you had this robot at home, and you were to leave for 2 weeks of holidays, would it be ok to leave the robot alone at home?''}. A point for anthropomorphism was counted, when the participant's answer indicated an anthropomorphic view of the robot. This process led to a ``qualitative score'' that reflects how far the robot is anthropomorphized. We used the following constructs to measure how far the robot is perceived as a human-like agent:

@@

Some of these questions were taken from previous work done by \cite{kahn_jr._robotic_2006} an \cite{weiss_i_2009}. In both studies, both questionnaires and an analysis of behavior were used to assess children's and adult's engagement with, and perception of the robot. This seems to be a promising approach to investigate anthropomorphism, as it is also possible to combine quantitative and qualitative techniques.

\subsection{Ethical aspects of anthropomorphism and human-like form of robots}
\label{sec:ethics}

@@

After all discussion, there remains the question how far a sustained anthropomorphic (human social) interaction between a human and a non-human can be favored, regarding ethical aspects. We need to rethink carefully the consequences that emotional bonds between humans and robots can have. They might be beneficial on one hand but can also hold negative implications. How far should a robot's design encourage us to lastingly project feelings or intentions to it? Why should a machine even react to us as if it had emotions? Can this be the reality we want to live in? What happens as soon as the artificial agent is considered a \textit{moral agent} \citep{sullins_when_2006}? Also, the use of \textit{socially interactive robots} in 
contexts involving elderly care or in rehabilitation and with autistic children is a controversially discussed topic 
\cite{robins_robots_2005}. Do we want robots to comfort us; robots that make us eventually even falling in love with them?
The anthropomorphic design of robots is critical. Why do we even need this? The HRI / robotics community needs to discuss these and other questions.

@@

@@

Based on our previous (and on-going) work, we hypothesized that there are qualitative differences and dynamics over time
in how humans relate to different types of robots. More concretely, we described
that people's perception of human-likeness in a robot and in turn their
human-social engagement with a robot is based on four main factors: (1) the (anthropomorphic) design of the robot, (2) individual characteristics of the human user, (3) the context of use / interaction, and (4) the interaction history (time factor). We further suggested
that people's tendency to anthropomorphize a robot is likely to change over time with
growing experience of the user with the system.

To formalize the variances in anthropomorphism, we
proposed a model that we call \textit{Dynamics of Anthropomorphism}.

 We introduced the concepts of
\emph{initial capital} and \emph{stabilized level of anthropomorphism} as compound
factors to characterize the profile of a given anthropomorphic (long-term) interaction.

The article also discussed the cognitive correlates of anthropomorphism. We
proposed three cognitive stages corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We showed how these stages relate to observable anthropomorphic
effects, and how they evolve over time.

@@

%With
%this initial model, we would like to extend the current understanding of
%anthropomorphism in robotics and human-robot interaction. This might in turn
%lead to a discussion about the short- and long-term benefits and other effects
%of the anthropomorphic design of robots and other artificial
%agents.
